---
title: "almond_profit_model"
format: html
---

```{r}
library(ggplot2)
library(tidyverse)
library(dplyr)
library(purrr)
```

preliminary research:
according to fruitgrowersupply.com (they report on industry standards), we found that average yield will range from 1400 - 2600 lbs/acre of almonds and will offer around 0.90 - 2.10 dollars. 

We decided to use 1,800 pounds / acre for the baseline almond yield in California (this was synthesis from the almond board of Calfornia.) When we convert this to tons/acre (as it is in the paper) we get 0.9 tons/acre. We decided to use $1.50 for our baseline price (in the middle of the range we found). We did some basic stoichiometry to make sure that our units all canceled out correctly, and we got 2,700 dollars / acre baseline price. 

From university of california agricultural issues center sample cost and returns studies, we found that the average cost of production over the past 30 years has been $4,500 / acre. 

We know that our final profit will be the (yield * unit price) - cost of production. We also want to call our yield function to account for the anamoly in yield. 

We have our function written, now we want to perform our informal insensitivity analysis on two parameters: `baseline_profit` and `production_cost`

```{r, warning=FALSE, message=FALSE}
# Create ranges for both baseline_profit and production_cost
baseline_profit_values <- seq(2000, 3500, by = 500)
production_cost_values <- seq(3500, 5500, by = 500)

# Create a data frame with all combinations of baseline_profit and production_cost
param_grid <- expand.grid(baseline_profit = baseline_profit_values, production_cost = production_cost_values)

# Run pmap to calculate profit for all combinations of baseline_profit and production_cost and make sure to include year
sensitivity_results <- pmap(
  param_grid,
  function(baseline_profit, production_cost) {
    # Calculate the yield anomaly and profit
    yield_anomaly_df <- profit(climate_df, baseline_profit = baseline_profit, production_cost = production_cost)
    
    # Create a new data frame with the combination of parameters and results, including year
    results_with_year <- yield_anomaly_df %>%
      select(wy, profit) %>%  # Select wy and profit
      mutate(baseline_profit = baseline_profit, production_cost = production_cost)  # Add parameters to the results
    
    # Return the results
    return(results_with_year)
  }
)

# Combine the list of data frames into a single data frame
sensitivity_results_df <- bind_rows(sensitivity_results)

# View the first few rows of the results to ensure it includes year and profit
head(sensitivity_results_df)
```

Plot that shows yield anomaly for each year, accounting for uncertainty in the parameters
```{r}
# Group by year and calculate mean + uncertainty (e.g., 2.5% and 97.5% quantiles)
profit_summary <- sensitivity_results_df %>%
  group_by(wy) %>%
  summarise(
    mean_profit = mean(profit),
    lower = quantile(profit, 0.025),
    upper = quantile(profit, 0.975)
  )

# Plot
ggplot(profit_summary, aes(x = wy, y = mean_profit)) +
  geom_ribbon(aes(ymin = lower, ymax = upper), fill = "pink") +
  geom_line(color = "blue", size = 1) +
  labs(
    title = "Annual Profit Over Time with Uncertainty",
    x = "Year",
    y = "Profit"
  )
```

```{r}
# Plot 2: Distribution of Profits Across All Years (like yield overview)

# All years together: show density with some jitter
ggplot(sensitivity_results_df, aes(x = profit)) +
  geom_density(fill = "blue", alpha = 0.5) +
  geom_rug(alpha = 0.3) +
  labs(
    title = "Distribution of Profit Across All Years and Scenarios",
    x = "Profit",
    y = "Density"
  ) +
  theme_minimal(base_size = 14)
```

There are a couple of things we can see from our plots: first, the annual profit over time, we see that 1995 was a really good year for their yield. We see from the first plot in the ribbons, that there is quite a large variability in the uncertainties. It seems like our model is pretty sensitive to uncertainty. The second plot shows us that the combinations of parameters most of the time give us a profit that is pretty low, around 2,000. There is a long tail which means it not as often, but sometimes reaches out to much higher profits. This aligns with what we saw in our previous plot with few large spikes. 
